### 2021-01-25 README.Rmd
First pass at refactoring README

```{r, eval = FALSE}
# install.packages("remotes")
# remotes::install_github("rudeboybert/forestecology")
```

@@

Here we provide an example with a small dataset.

@@

First we combine two example census data files into a single `tibble` and compute the growth of all surviving individuals.

```{r}
# Read in census files
data(census_df1_ex, census_df2_ex)

@@

# Name of variable in tibble that uniquely identifies each stem:
id <- "ID"

# Merge both censuses and compute growth:
ex_growth_df <-
  compute_growth(census_df1_ex, census_df2_ex_no_r, id) %>%
  mutate(
    sp = to_any_case(sp),
    # Must be a factor
    sp = as.factor(sp)
  )
```

All growth models in the package assume that two individuals compete if they are less than a specified distance away. Here we set this distance in `max_dist` and put all individuals within this distance of a plot boundary in a buffer (since we do not have full information about their competitors).

```{r}
# Load study region:
data("ex_study_region")

# Set max dist
max_dist <- 1

# Add buffer
ex_growth_df <- ex_growth_df %>%
  add_buffer_variable(direction = "in", size = max_dist, region = ex_study_region)

ggplot() +
  geom_sf(data = ex_growth_df, aes(col = buffer), size = 2)
```

We then set blocks/folds for the spatial cross-validation scheme. We this both manually and using the [`blockCV`](https://github.com/rvalavi/blockCV) package. Note that here we define the geometric structure of the folds manually using `sfheaders::sf_polygon()`:

```{r}
fold1 <- rbind(c(0, 0), c(5, 0), c(5, 5), c(0, 5), c(0, 0))
fold2 <- rbind(c(5, 0), c(10, 0), c(10, 5), c(5, 5), c(5, 0))
blocks <- bind_rows(
  sf_polygon(fold1),
  sf_polygon(fold2)
) %>%
  mutate(foldID = c(1, 2))
```

Next we assign each point to folds using `blockCV::spatialBlock()`

```{r}
ex_cv_grid <- spatialBlock(
  speciesData = ex_growth_df,
  verbose = FALSE,
  k = 2,
  selection = "systematic",
  blocks = blocks,
  showBlocks = FALSE
)

# Add foldID to data
ex_growth_df <- ex_growth_df %>%
  mutate(foldID = ex_cv_grid$foldID %>% as.factor())

# Deliverable
ggplot() +
  geom_sf(data = ex_growth_df, aes(col = buffer, shape = foldID), size = 2) +
  geom_sf(data = blocks, fill = "transparent")

# Convert to sf object
ex_cv_grid_sf <- ex_cv_grid$blocks %>%
  st_as_sf()
```

We then create a `focal_v_comp` tibble which has a row for each competing pair of individuals (each pair of individuals within `max_dist` of one another):

```{r}
focal_vs_comp_ex <- ex_growth_df %>%
  create_focal_vs_comp(max_dist, cv_grid_sf = ex_cv_grid_sf, id = "ID")
```

With `focal_v_comp`, we can then run the growth model and get predictions from it. Here we run the model without cross validation. Posterior estimates can be plotted to give, for example, the lambda matrix of competition coefficients.

```{r}
# Fit model
comp_bayes_lm_ex <- focal_vs_comp_ex %>%
  comp_bayes_lm(prior_param = NULL, run_shuffle = FALSE)

# Get predicted dbh values values
predictions <- focal_vs_comp_ex %>%
  mutate(growth_hat = predict(comp_bayes_lm_ex, focal_vs_comp_ex))

# Compute RMSE of true vs predicted dbh values
predictions %>%
  rmse(truth = growth, estimate = growth_hat) %>%
  pull(.estimate)

# Plot posteriors
comp_bayes_lm_ex %>%
  autoplot(type = "intercepts")
comp_bayes_lm_ex %>%
  autoplot(type = "dbh_slopes")
comp_bayes_lm_ex %>%
  autoplot(type = "competition")
```

Here we repeat the process but with cross-validation. Note the increase in RMSE, reflecting the fact that our original estimate of model error was overly optimistic.

```{r}
# Fit model with cross-validation
ex_bw <- focal_vs_comp_ex %>%
  run_cv(max_dist = max_dist, cv_grid = ex_cv_grid_sf) %>%
  right_join(ex_growth_df, by = c("focal_ID" = "ID"))

# Compute RMSE of true vs predicted dbh values
ex_bw %>%
  rmse(truth = growth, estimate = growth_hat) %>%
  pull(.estimate)
```

For fuller examples on actual datasets see ___.

>>>>

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("rudeboybert/forestecology")
```

@@

We present an example analysis using toy data pre-loaded into the package.

@@

### Preprocess census data

We first combine data from two repeat censuses into a single `tibble` and compute the growth of all surviving individuals. Note that per Allen and Kim (2020) we first remove resprouts from the second census.

```{r}
data(census_df1_ex, census_df2_ex)

@@

# Merge both censuses and compute growth. Note we require any species variable
# to be coded as a factor
census_combined_ex <-
  compute_growth(census_df1_ex, census_df2_ex_no_r, id = "ID") %>%
  mutate(sp = to_any_case(sp) %>% as.factor())
```

### Add spatial information

Our growth model assumes that two individual trees compete if they are less than a pre-specified distance `max_dist` apart. Furthermore, we define a buffer region of size `max_dist` from the boundary of the study region. All trees outside this buffer region will be our "focal" trees of interest since we have complete competitor information on them. All trees inside this buffer region will only be considered as "competitor" trees to "focal" trees.

```{r}
data("ex_study_region")

# Set competitor distance
max_dist <- 1

# Create sf representation of buffer region
buffer_region <- ex_study_region %>% 
  compute_buffer_region(direction = "in", size = max_dist)

# Add buffer variable
census_combined_ex <- census_combined_ex %>%
  add_buffer_variable(direction = "in", size = max_dist, region = ex_study_region)

base_plot <- ggplot() +
  geom_sf(data = ex_study_region, fill = "transparent") +
  geom_sf(data = buffer_region, fill = "transparent", linetype = "dashed")

base_plot + 
  geom_sf(data = census_combined_ex, aes(col = buffer), size = 2)
```

We then manually define the blocks/folds for our spatial cross-validation scheme and convert them to an `sf` object using `sfheaders::sf_polygon()`:

```{r}
fold1 <- rbind(c(0, 0), c(5, 0), c(5, 5), c(0, 5), c(0, 0))
fold2 <- rbind(c(5, 0), c(10, 0), c(10, 5), c(5, 5), c(5, 0))

blocks <- bind_rows(
  sf_polygon(fold1),
  sf_polygon(fold2)
) %>%
  mutate(foldID = c(1, 2))
```

Next we assign each tree to blocks/folds using the `spatialBlock()` function from the [`blockCV`](https://github.com/rvalavi/blockCV) package.

```{r}
ex_cv_grid <- spatialBlock(
  speciesData = census_combined_ex, k = 2, selection = "systematic", blocks = blocks,
  showBlocks = FALSE, verbose = FALSE
)

# Add fold information to data
census_combined_ex <- census_combined_ex %>%
  mutate(foldID = ex_cv_grid$foldID %>% factor())

base_plot + 
  geom_sf(data = census_combined_ex, aes(col = buffer, shape = foldID), size = 2) +
  geom_sf(data = blocks, fill = "transparent", col = "orange")
```

We also extract the spatial cross-validation grid itself as an `sf` object for later use.

```{r}
ex_cv_grid_sf <- ex_cv_grid$blocks %>%
  st_as_sf()

ggplot() +
  geom_sf(data = ex_cv_grid_sf)
```

### Define focal versus competitor trees

We then create a list-column data frame (via `tidyr::nest()`) where each row represents a focal tree of interest and the `comp` variable contains nested information on each of it's competitors

which has a row for each competing pair of individuals (each pair of individuals within `max_dist` of one another):

```{r}
focal_vs_comp_ex <- census_combined_ex %>%
  create_focal_vs_comp(max_dist, cv_grid_sf = ex_cv_grid_sf, id = "ID")

# Nested view
focal_vs_comp_ex

# Unnested view:
focal_vs_comp_ex %>% 
  unnest(cols = "comp")
```

### Fit model and make predictions

We then fit our competitor growth model as specified in Allen and Kim (2020)

```{r}
comp_bayes_lm_ex <- focal_vs_comp_ex %>%
  comp_bayes_lm(prior_param = NULL, run_shuffle = FALSE)
```

Since the resulting output is an S3 object of class `comp_bayes_lm`, we can both print and visualize its output: the posterior distribution of all the linear regression parameters.

```{r}
# Object summary
comp_bayes_lm_ex

# Visualizations
autoplot(comp_bayes_lm_ex, type = "intercepts")
autoplot(comp_bayes_lm_ex, type = "dbh_slopes")
autoplot(comp_bayes_lm_ex, type = "competition")
```

Furthermore, we append fitted/predicted growth values for each tree using `predict.comp_bayes_lm()` to our data frame and then compute the root mean squared error (RMSE).

```{r}
focal_vs_comp_ex <- focal_vs_comp_ex %>%
  mutate(growth_hat = predict(comp_bayes_lm_ex, focal_vs_comp_ex))

# Compute RMSE
focal_vs_comp_ex %>%
  yardstick::rmse(truth = growth, estimate = growth_hat) %>%
  pull(.estimate)
```

### Run spatial cross-validation

Here we repeat the process but with running spatial cross-validation: we fit the data to one fold and apply it to the other. Note the increase in RMSE, reflecting the fact that our original estimate of model error was overly optimistic as it did not account for spatial autocorrelation.

```{r}
ex_bw <- focal_vs_comp_ex %>%
  run_cv(max_dist = max_dist, cv_grid = ex_cv_grid_sf) %>%
  right_join(census_combined_ex, by = c("focal_ID" = "ID"))

# Compute RMSE of spatially cross-validated model
ex_bw %>%
  rmse(truth = growth, estimate = growth_hat) %>%
  pull(.estimate)
```

<!--For fuller examples on actual datasets see ___.-->
